---
title: "Unit testing movement functions"
author: "Simon Mills"
date: "13/10/2020"
output: html_document
---
# Overview
__Series of tests on simplified matrices to check function behaviour is as expected.__

Movement functions all depend on igraph for constructing network of linked cells and then various approaches to quantifying  movement through these. Broad building blocks are:

(1) take a matrix of forest cover and elevation (though could in principle be any matrix of permissible movement and any gradient that you want directional movement along).
(2) Identify permitted cells within a given search radius (degree of gap-crossing)
(3) Trim all adjacencies to those that are along or up-slope (along-slope movement will be by definition bidirectional, while upslope is unidirectional)
(4) Build graph
(5) With graph in hand, generate statistics of interest 

Notes: 

- The full raster has to have a margin of "not-permitted" (i.e. 0-valued cells), otherwise it will wrap around edges. Have a function to do this, but currently do the matrix processing outside of the function: may be safest to either wrap this in a helper function, or stick it in the graph function directly.. 

- Method is pretty fast: using Cordillera Oriental de Colombia at 100m resolution, can create graph for 48 million cells in 2.5 minutes, and this appears to be the main time overhead with subsequent functions (i.e. 5) being negligible relative to this (e.g. 10 seconds). Time and memory should (in principle) scale approximately linearly, so plan would be to drop it to 60m resolution and try that. Ultimately think it should go to 30m. 

```{r setup, include=FALSE}
library(raster); library(igraph); library(dplyr)
source("functions_igraph.R")

# rasters index cells horizontally rather than vertically, which will mess up code
# written for matrices, but plotting them without rotation and safely (by reducing
# resolution for large matrices is useful), so use this. 
plot_mat <- function(x, ...) plot(raster(x), ...)

```

# Test 1: spread of algorithm under varying gap-crossing
Note this is just isolating the mechanism of spread through the graph, with varying gap-crossing ability. This spread function is not called directly in the main functions. 

```{r}
# read 65*50 matrix that has "permitted movement" cells spaced contiguously, 
# 1-cell removed, and 2-cell removed
test_mat <- as.matrix(raster("../data/test_imgs/test_gap_crossing_rule.png"))
start_cell <- which(test_mat[] == 237)
test_mat[test_mat[] %in% c(0, 237)] <- 1
test_mat[test_mat[] == 255] <- 0

# for spread function
n_row <- dim(test_mat)[1]
n_col <- dim(test_mat)[2]
permitted_cells <- which(test_mat == 1)

# run
fill1 <- spread_igraph(permitted_cells, start_cell, n_row, n_col, rule=1)
fill2 <- spread_igraph(permitted_cells, start_cell, n_row, n_col, rule=2)
fill3 <- spread_igraph(permitted_cells, start_cell, n_row, n_col, rule=3)

# fill points reached by algorithm
mat1 <- test_mat; mat1[fill1] <- 2
mat2 <- test_mat; mat2[fill2] <- 2
mat3 <- test_mat; mat3[fill3] <- 2

# save results
par(mfrow=c(2,2))
permitted_plot <- test_mat
permitted_plot[start_cell] <- 2
plot_mat(permitted_plot, asp=1, main = "Permitted movement & start")
plot_mat(mat1, asp=1, main = "Spread, 1-cell search radius")
plot_mat(mat2, asp=1, main = "Spread, 2-cell search radius")
plot_mat(mat3, asp=1, main = "Spread, 3-cell search radius")
```

```{r}
# read 65*50 matrix that has "permitted movement" cells spaced contiguously, 
# 1-cell removed, and 2-cell removed
test_mat <- as.matrix(raster("../data/test_imgs/pikachu.png"))
start_cell <- which(test_mat[] == 237)
test_mat[test_mat[] %in% c(5, 237)] <- 1
test_mat[test_mat[] > 1] <-0

# for spread function
n_row <- dim(test_mat)[1]
n_col <- dim(test_mat)[2]
permitted_cells <- which(test_mat == 1)

# run
fill1 <- spread_igraph(permitted_cells, start_cell, n_row, n_col, rule=1)
fill2 <- spread_igraph(permitted_cells, start_cell, n_row, n_col, rule=2)
fill3 <- spread_igraph(permitted_cells, start_cell, n_row, n_col, rule=3)

# fill points reached by algorithm
mat1 <- test_mat; mat1[fill1] <- 20
mat2 <- test_mat; mat2[fill2] <- 20
mat3 <- test_mat; mat3[fill3] <- 20

# save results
par(mfrow=c(2,2))
permitted_plot <- test_mat
permitted_plot[start_cell] <- 20
plot_mat(permitted_plot, asp=1, main = "Permitted movement & start")
plot_mat(mat1, asp=1, main = "Spread, 1-cell search radius")
plot_mat(mat2, asp=1, main = "Spread, 2-cell search radius")
plot_mat(mat3, asp=1, main = "Spread, 3-cell search radius")
```

# Test 2: get all cells up-gradient from a start cell
(i.e. travelling from smaller values to larger, as is the case moving up or along an elevational gradient)
```{r}
# get all cells that are accessible up-gradient from a start-cell
test_mat <- as.matrix(raster("../figures/test_range_shift_1.png"))

# get start cell
in_range <- which(test_mat[] == 237)
# convert start range and permitted movements to 1
test_mat[test_mat[] %in% c(0, 237)] <- 1
# convert not-permitted to 0
test_mat[test_mat[] == 255] <- 0

# for passing to function
permitted_cells <- which(test_mat[] == 1)
n_row <- dim(test_mat)[1]
n_col <- dim(test_mat)[2]

# generate gradient
df_gradient <- expand.grid(x = 1:n_col, y = 1:n_row) %>%
    arrange(x, y) %>%
    mutate(xy = 20-x) 
# for passing to simulation
ele_vec <- df_gradient$xy

# run graph and ranges
graph <- generate_graph(ele_vec, permitted_cells, n_row, n_col, rule=1)
ranges <- get_range(graph, ele_vec, start_cells = in_range)

# visualise
par(mfrow=c(2,2))
mat_tmp <- test_mat; mat_tmp[in_range] <- 2
plot_mat(mat_tmp, main="Permissible movement & start-range (green)")
mat_tmp <- test_mat; mat_tmp[] <- ele_vec
plot_mat(mat_tmp, main="Gradient")
mat_tmp <- test_mat; mat_tmp[ranges[!is.na(ordering), id_cell]] <- 2
plot_mat(mat_tmp, main="Accessed cells")
```

## Variant with connection upslope 
```{r}
# get all cells that are accessible up-gradient from a start-cell (alternative
# version with linkage upslope)
test_mat <- as.matrix(raster("../figures/test_range_shift_2.png"))

# get start cell
in_range <- which(test_mat[] == 237)
# convert start range and permitted movements to 1
test_mat[test_mat[] %in% c(0, 237)] <- 1
# convert not-permitted to 0
test_mat[test_mat[] == 255] <- 0

# for passing to function
permitted_cells <- which(test_mat[] == 1)
n_row <- dim(test_mat)[1]
n_col <- dim(test_mat)[2]

# generate gradient
df_gradient <- expand.grid(x = 1:n_col, y = 1:n_row) %>%
    arrange(x, y) %>%
    mutate(xy = 20-x) 
# for passing to simulation
ele_vec <- df_gradient$xy

# run graph and ranges
graph <- generate_graph(ele_vec, permitted_cells, n_row, n_col, rule=1)
ranges <- get_range(graph, ele_vec, start_cells = in_range)

# visualise
par(mfrow=c(2,2))
mat_tmp <- test_mat; mat_tmp[in_range] <- 2
plot_mat(mat_tmp, main="Permissible movement & start-range (green)")
mat_tmp <- test_mat; mat_tmp[] <- ele_vec
plot_mat(mat_tmp, main="Gradient")
mat_tmp <- test_mat; mat_tmp[ranges[!is.na(ordering), id_cell]] <- 2
plot_mat(mat_tmp, main="Accessed cells")
```

## Variant with up-gradient movement from multiple start-points
```{r}
test_mat <- as.matrix(raster("../figures/test_range_shift_3.png"))

# get start cell
in_range <- which(test_mat[] == 237)
# convert start range and permitted movements to 1
test_mat[test_mat[] %in% c(0, 237)] <- 1
# convert not-permitted to 0
test_mat[test_mat[] == 255] <- 0

# for passing to function
permitted_cells <- which(test_mat[] == 1)
n_row <- dim(test_mat)[1]
n_col <- dim(test_mat)[2]

# generate gradient
df_gradient <- expand.grid(x = 1:n_col, y = 1:n_row) %>%
    arrange(x, y) %>%
    mutate(xy = 20-x) 
# for passing to simulation
ele_vec <- df_gradient$xy

# run graph and ranges
graph <- generate_graph(ele_vec, permitted_cells, n_row, n_col, rule=1)
ranges <- get_range(graph, ele_vec, start_cells = in_range)

# visualise
par(mfrow=c(2,2))
mat_tmp <- test_mat; mat_tmp[in_range] <- 2
plot_mat(mat_tmp, main="Permissible movement & start-range (green)")
mat_tmp <- test_mat; mat_tmp[] <- ele_vec
plot_mat(mat_tmp, main="Gradient")
mat_tmp <- test_mat; mat_tmp[ranges[!is.na(ordering), id_cell]] <- 2
plot_mat(mat_tmp, main="Accessed cells")
```

# Test X (out of action; eval=F)
```{r, eval=F}
# Does function correctly track niche through time?
# create an elevational gradient
df_gradient <- expand.grid(x = 1:n_col, y = 1:n_row) %>%
    mutate(xy = (x*y), xy_scale = (xy-min(xy))/max(xy-min(xy)) * 1000) %>%
    arrange(x, y)

# for passing to simulation
ele_vec <- df_gradient$xy_scale
start_cells <- min(which(test_mat[] == 1))
permitted_cells <- which(permitted_movement == 1)
n_iter <- 300
lwr_vec <- 0 + 0:(n_iter+1)
upr_vec <- 100 + 0:(n_iter+1)

# simulate
ranges_sim <- sim_new_range_igraph(n_iter, ele_vec, lwr_vec, upr_vec,
                                   permitted_cells, start_cells, n_row, n_col, rule=2)

# save
png("figures/test_rangesim_rule2.png", height=200, width=110, units="mm", res=100)
par(mfrow = c(4,2))
plot_mat(matrix(ele_vec, ncol=n_col, nrow=n_row), main="Elevation gradient")
mat_2 <- test_mat; mat_2[mat_2 == 1] <- ele_vec[mat_2==1]
plot_mat(mat_2, main="Permitted movement & gradient")
# plot results
mat1 <- test_mat; mat1[ranges_sim[[1]]] <- 20
plot_mat(mat1, main="t[1]")
mat1 <- test_mat; mat1[ranges_sim[[50]]] <- 20
plot_mat(mat1, main="t[50]")
mat1 <- test_mat; mat1[ranges_sim[[100]]] <- 20
plot_mat(mat1, main="t[100]")
mat1 <- test_mat; mat1[ranges_sim[[150]]] <- 20
plot_mat(mat1, main="t[150]")
mat1 <- test_mat; mat1[ranges_sim[[200]]] <- 20
plot_mat(mat1, main="t[200]")
mat1 <- test_mat; mat1[ranges_sim[[250]]] <- 20
plot_mat(mat1, main="t[250]")
dev.off()

# generate gif
# library(animation)
# saveHTML({
#     for(i in seq(1, n_iter, round(n_iter/50, 0))) {
#         mat_t <- test_mat; mat_t[ranges_sim[[i]]] <- 10
#         plot_mat(mat_t)
#     }
# },title = "Simulated spread, rule=1", description = desc, verbose = FALSE)

## getting connectivity measures
# permitted_cells <- which(permitted_vec == 1)
# rule <- 1
# dat <- data.frame(x1 = rep(permitted_cells, (rule*2 + 1)^2 - 1),
#                   x2 = get_adjacent(permitted_cells, n_row, n_col, rule)) %>%
#     filter(x2 %in% permitted_cells)
# graph <- graph_from_data_frame(dat)
# members <- components(graph)$membership %>%
#     data.frame(membership = ., id_cell = as.integer(names(.)))
# members$id_cell[members$membership %in% members$membership[members$id_cell %in% start_cells]]
```
