---
title: "Description and unit testing movement functions"
author: "Simon Mills"
date: "13/10/2020"
output: 
    html_document:
        number_sections: true
---
# Overview
__Series of tests on simplified matrices to check function behaviour is as expected.__

Movement functions all depend on igraph for constructing network of linked cells and then various approaches to quantifying  movement through these. Broad building blocks are:

(1) take a matrix of forest cover and elevation (though could in principle be any matrix of permissible movement and any gradient that you want directional movement along).
(2) Identify permitted cells within a given search radius (degree of gap-crossing)
(3) Trim all adjacencies to those that are along or up-slope (along-slope movement will be by definition bidirectional, while upslope is unidirectional)
(4) Build graph
(5) With graph in hand, generate statistics of interest 

## Important:

- There is a bug in the igraph function that will very (very) infrequently generate duplicate vertices for
*some* values by generating scientific notation equivalents (eg. a vertex for both 500000 *and* 5e+05). So far have observed x values of of 5, 7, 9, and 10. In all cases observed so far, the xe+05 version only has incoming edges, while the x00000 version only has outgoing edges. 

- It apparently only affects a tiny number of cases (e.g. 4 duplicated nodes in a network of ~2.7e06 vertices), so is going to have an ignorable impact on results, but have fixed regardless. 

- UPDATE: For Cordillera Central de Ecuador, 24 vertices are duplicated in this manner (presumably stored as e+06 and e+07), again this is a diminishingly small number out of a graph of 28 million vertices. 
```{r, eval=F}
[1] 100000000  64000000  73000000  88000000  77000000  65000000 112000000  61000000  49000000  66000000  98000000
[12]  89000000  38000000 126000000 113000000  80000000  60000000 102000000 103000000  23000000 114000000  59000000
[23]  90000000  71000000
```

- The workaround is to identify these failure cases, reroute the incoming edges to the x000.. version (creating a correctly connected vertex in the graph), and then remove the xe+0.. version of the vertices. This appears to resolve the issue. 

- The sanity check is that there should be the same number of cells in the eventuating graph and set of cells identified as disconnected from the graph as there are in the total number of permitted cells provided at the outset. I've also stored the list of duplicated cells to confirm that there is no additional behaviour beyond that already described here.  

- Need to generate a reproducible example and submit as an error.

## Notes: 

- The full raster has to have a margin of "not-permitted" (i.e. 0-valued cells), otherwise it will wrap around edges. Have a function to do this, but currently do the matrix processing outside of the function: may be safest to either wrap this in a helper function, or stick it in the graph function directly.. 

- Method is pretty fast: using Cordillera Oriental de Colombia at 100m resolution, can create graph for 48 million cells in 2.5 minutes, using the Cordillera Central de Ecuador at 60m resolution (137 million cells) takes 14.5 minutes. This appears to be the main time overhead with subsequent functions being negligible relative to this, though will run them many times, which can make these latter steps similar in terms of time. 

- Memory usage can be substantial: Cordillera Central de Ecuador will run locally (just) but appears to use about 24GB to build the graph (though the final object is only 6GB). Saving this as a binary file (RDS), is slow (~approx 5-10 minutes, but to read is only ~100s)

- Time and memory should scale approximately linearly (and this does indeed appear to be the case). Cordillera Central de Ecuador is a large mountain range, and the results from this therefore suggest that *any* tropical mountain range should be computationally feasable at this spatial resolution. 

----
# UNIT TESTS
```{r setup, message=F}
library(raster); library(igraph); library(dplyr); library(data.table)
source("functions_igraph.R")

# rasters index cells horizontally rather than vertically, which will mess up code
# written for matrices, but plotting them without rotation and safely (by reducing
# resolution for large matrices is useful), so use this. 
plot_mat <- function(x, ...) plot(raster(x), ...)

```


## Spread of algorithm under varying gap-crossing
Note this is just isolating the mechanism of spread through the graph, with varying gap-crossing ability. This spread function is not called directly in the main functions. 

```{r}
# read 65*50 matrix that has "permitted movement" cells spaced contiguously, 
# 1-cell removed, and 2-cell removed
test_mat <- as.matrix(raster("../data/test_imgs/test_gap_crossing_rule.png"))
start_cell <- which(test_mat[] == 237)
test_mat[test_mat[] %in% c(0, 237)] <- 1
test_mat[test_mat[] == 255] <- 0

# for spread function
n_row <- dim(test_mat)[1]
n_col <- dim(test_mat)[2]
permitted_cells <- which(test_mat == 1)

# run
fill1 <- spread_igraph(permitted_cells, start_cell, n_row, n_col, rule=1)
fill2 <- spread_igraph(permitted_cells, start_cell, n_row, n_col, rule=2)
fill3 <- spread_igraph(permitted_cells, start_cell, n_row, n_col, rule=3)

# fill points reached by algorithm
mat1 <- test_mat; mat1[fill1] <- 2
mat2 <- test_mat; mat2[fill2] <- 2
mat3 <- test_mat; mat3[fill3] <- 2

# save results
par(mfrow=c(2,2))
permitted_plot <- test_mat
permitted_plot[start_cell] <- 2
plot_mat(permitted_plot, asp=1, main = "Permitted movement & start")
plot_mat(mat1, asp=1, main = "Spread, 1-cell search radius")
plot_mat(mat2, asp=1, main = "Spread, 2-cell search radius")
plot_mat(mat3, asp=1, main = "Spread, 3-cell search radius")
```

```{r}
# read 65*50 matrix that has "permitted movement" cells spaced contiguously, 
# 1-cell removed, and 2-cell removed
test_mat <- as.matrix(raster("../data/test_imgs/pikachu.png"))
start_cell <- which(test_mat[] == 237)
test_mat[test_mat[] %in% c(5, 237)] <- 1
test_mat[test_mat[] > 1] <-0

# for spread function
n_row <- dim(test_mat)[1]
n_col <- dim(test_mat)[2]
permitted_cells <- which(test_mat == 1)

# run
fill1 <- spread_igraph(permitted_cells, start_cell, n_row, n_col, rule=1)
fill2 <- spread_igraph(permitted_cells, start_cell, n_row, n_col, rule=2)
fill3 <- spread_igraph(permitted_cells, start_cell, n_row, n_col, rule=3)

# fill points reached by algorithm
mat1 <- test_mat; mat1[fill1] <- 20
mat2 <- test_mat; mat2[fill2] <- 20
mat3 <- test_mat; mat3[fill3] <- 20

# save results
par(mfrow=c(2,2))
permitted_plot <- test_mat
permitted_plot[start_cell] <- 20
plot_mat(permitted_plot, asp=1, main = "Permitted movement & start")
plot_mat(mat1, asp=1, main = "Spread, 1-cell search radius")
plot_mat(mat2, asp=1, main = "Spread, 2-cell search radius")
plot_mat(mat3, asp=1, main = "Spread, 3-cell search radius")
```

## Get all cells up-gradient from a start cell
(i.e. travelling from smaller values to larger, as is the case moving up or along an elevational gradient)
```{r}
# get all cells that are accessible up-gradient from a start-cell
test_mat <- as.matrix(raster("../figures/test_range_shift_1.png"))

# get start cell
in_range <- which(test_mat[] == 237)
# convert start range and permitted movements to 1
test_mat[test_mat[] %in% c(0, 237)] <- 1
# convert not-permitted to 0
test_mat[test_mat[] == 255] <- 0

# for passing to function
permitted_cells <- which(test_mat[] == 1)
n_row <- dim(test_mat)[1]
n_col <- dim(test_mat)[2]

# generate gradient
df_gradient <- expand.grid(x = 1:n_col, y = 1:n_row) %>%
    arrange(x, y) %>%
    mutate(xy = 20-x) 
# for passing to simulation
ele_vec <- df_gradient$xy

# run graph and ranges
graph <- generate_graph(ele_vec, permitted_cells, n_row, n_col, rule=1)
ranges <- get_range(graph, ele_vec, start_cells = in_range)

# visualise
par(mfrow=c(2,2))
mat_tmp <- test_mat; mat_tmp[in_range] <- 2
plot_mat(mat_tmp, main="Permissible movement & start-range (green)")
mat_tmp <- test_mat; mat_tmp[] <- ele_vec
plot_mat(mat_tmp, main="Gradient")
mat_tmp <- test_mat; mat_tmp[ranges[!is.na(ordering), id_cell]] <- 2
plot_mat(mat_tmp, main="Accessed cells")
```

### Variant with connection upslope 
```{r}
# get all cells that are accessible up-gradient from a start-cell (alternative
# version with linkage upslope)
test_mat <- as.matrix(raster("../figures/test_range_shift_2.png"))

# get start cell
in_range <- which(test_mat[] == 237)
# convert start range and permitted movements to 1
test_mat[test_mat[] %in% c(0, 237)] <- 1
# convert not-permitted to 0
test_mat[test_mat[] == 255] <- 0

# for passing to function
permitted_cells <- which(test_mat[] == 1)
n_row <- dim(test_mat)[1]
n_col <- dim(test_mat)[2]

# generate gradient
df_gradient <- expand.grid(x = 1:n_col, y = 1:n_row) %>%
    arrange(x, y) %>%
    mutate(xy = 20-x) 
# for passing to simulation
ele_vec <- df_gradient$xy

# run graph and ranges
graph <- generate_graph(ele_vec, permitted_cells, n_row, n_col, rule=1)
ranges <- get_range(graph, ele_vec, start_cells = in_range)

# visualise
par(mfrow=c(2,2))
mat_tmp <- test_mat; mat_tmp[in_range] <- 2
plot_mat(mat_tmp, main="Permissible movement & start-range (green)")
mat_tmp <- test_mat; mat_tmp[] <- ele_vec
plot_mat(mat_tmp, main="Gradient")
mat_tmp <- test_mat; mat_tmp[ranges[!is.na(ordering), id_cell]] <- 2
plot_mat(mat_tmp, main="Accessed cells")
```

### Variant with connection upslope (swapped gradient)
```{r}
# get all cells that are accessible up-gradient from a start-cell (alternative
# version with linkage upslope)
test_mat <- as.matrix(raster("../figures/test_range_shift_2.png"))

# get start cell
in_range <- which(test_mat[] == 237)
# convert start range and permitted movements to 1
test_mat[test_mat[] %in% c(0, 237)] <- 1
# convert not-permitted to 0
test_mat[test_mat[] == 255] <- 0

# for passing to function
permitted_cells <- which(test_mat[] == 1)
n_row <- dim(test_mat)[1]
n_col <- dim(test_mat)[2]

# generate gradient
df_gradient <- expand.grid(x = 1:n_col, y = 1:n_row) %>%
    arrange(x, y) %>%
    mutate(xy = x) 
# for passing to simulation
ele_vec <- df_gradient$xy

# run graph and ranges
graph <- generate_graph(ele_vec, permitted_cells, n_row, n_col, rule=1)
ranges <- get_range(graph, ele_vec, start_cells = in_range)

# visualise
par(mfrow=c(2,2))
mat_tmp <- test_mat; mat_tmp[in_range] <- 2
plot_mat(mat_tmp, main="Permissible movement & start-range (green)")
mat_tmp <- test_mat; mat_tmp[] <- ele_vec
plot_mat(mat_tmp, main="Gradient")
mat_tmp <- test_mat; mat_tmp[ranges[!is.na(ordering), id_cell]] <- 2
plot_mat(mat_tmp, main="Accessed cells")
```

### Variant with up-gradient movement from multiple start-points
```{r}
test_mat <- as.matrix(raster("../figures/test_range_shift_3.png"))

# get start cell
in_range <- which(test_mat[] == 237)
# convert start range and permitted movements to 1
test_mat[test_mat[] %in% c(0, 237)] <- 1
# convert not-permitted to 0
test_mat[test_mat[] == 255] <- 0

# for passing to function
permitted_cells <- which(test_mat[] == 1)
n_row <- dim(test_mat)[1]
n_col <- dim(test_mat)[2]

# generate gradient
df_gradient <- expand.grid(x = 1:n_col, y = 1:n_row) %>%
    arrange(x, y) %>%
    mutate(xy = 20-x) 
# for passing to simulation
ele_vec <- df_gradient$xy

# run graph and ranges
graph <- generate_graph(ele_vec, permitted_cells, n_row, n_col, rule=1)
ranges <- get_range(graph, ele_vec, start_cells = in_range)

# visualise
par(mfrow=c(2,2))
mat_tmp <- test_mat; mat_tmp[in_range] <- 2
plot_mat(mat_tmp, main="Permissible movement & start-range (green)")
mat_tmp <- test_mat; mat_tmp[] <- ele_vec
plot_mat(mat_tmp, main="Gradient")
mat_tmp <- test_mat; mat_tmp[] <- NA; mat_tmp[ranges[!is.na(ordering), id_cell]] <- 2
mat_tmp[ranges[start_cell == T, id_cell]] <- 4
plot_mat(mat_tmp, main="Accessed cells and start-cell")
```

### Variant with up-gradient movement from multiple start-points (swapped gradient)
```{r}
test_mat <- as.matrix(raster("../figures/test_range_shift_3.png"))

# get start cell
in_range <- which(test_mat[] == 237)
# convert start range and permitted movements to 1
test_mat[test_mat[] %in% c(0, 237)] <- 1
# convert not-permitted to 0
test_mat[test_mat[] == 255] <- 0

# for passing to function
permitted_cells <- which(test_mat[] == 1)
n_row <- dim(test_mat)[1]
n_col <- dim(test_mat)[2]

# generate gradient
df_gradient <- expand.grid(x = 1:n_col, y = 1:n_row) %>%
    arrange(x, y) %>%
    mutate(xy = x) 
# for passing to simulation
ele_vec <- df_gradient$xy

# run graph and ranges
graph <- generate_graph(ele_vec, permitted_cells, n_row, n_col, rule=1)
ranges <- get_range(graph, ele_vec, start_cells = in_range)

# visualise
par(mfrow=c(2,2))
mat_tmp <- test_mat; mat_tmp[in_range] <- 2
plot_mat(mat_tmp, main="Permissible movement & start-range (green)")
mat_tmp <- test_mat; mat_tmp[] <- ele_vec
plot_mat(mat_tmp, main="Gradient")
mat_tmp <- test_mat; mat_tmp[] <- NA; mat_tmp[ranges[!is.na(ordering), id_cell]] <- 2
mat_tmp[ranges[start_cell == T, id_cell]] <- 4
plot_mat(mat_tmp, main="Accessed cells and start-cell")

```

## Test X (out of action; eval=F)
```{r, eval=F}
# Does function correctly track niche through time?
# create an elevational gradient
df_gradient <- expand.grid(x = 1:n_col, y = 1:n_row) %>%
    mutate(xy = (x*y), xy_scale = (xy-min(xy))/max(xy-min(xy)) * 1000) %>%
    arrange(x, y)

# for passing to simulation
ele_vec <- df_gradient$xy_scale
start_cells <- min(which(test_mat[] == 1))
permitted_cells <- which(permitted_movement == 1)
n_iter <- 300
lwr_vec <- 0 + 0:(n_iter+1)
upr_vec <- 100 + 0:(n_iter+1)

# simulate
ranges_sim <- sim_new_range_igraph(n_iter, ele_vec, lwr_vec, upr_vec,
                                   permitted_cells, start_cells, n_row, n_col, rule=2)

# save
png("figures/test_rangesim_rule2.png", height=200, width=110, units="mm", res=100)
par(mfrow = c(4,2))
plot_mat(matrix(ele_vec, ncol=n_col, nrow=n_row), main="Elevation gradient")
mat_2 <- test_mat; mat_2[mat_2 == 1] <- ele_vec[mat_2==1]
plot_mat(mat_2, main="Permitted movement & gradient")
# plot results
mat1 <- test_mat; mat1[ranges_sim[[1]]] <- 20
plot_mat(mat1, main="t[1]")
mat1 <- test_mat; mat1[ranges_sim[[50]]] <- 20
plot_mat(mat1, main="t[50]")
mat1 <- test_mat; mat1[ranges_sim[[100]]] <- 20
plot_mat(mat1, main="t[100]")
mat1 <- test_mat; mat1[ranges_sim[[150]]] <- 20
plot_mat(mat1, main="t[150]")
mat1 <- test_mat; mat1[ranges_sim[[200]]] <- 20
plot_mat(mat1, main="t[200]")
mat1 <- test_mat; mat1[ranges_sim[[250]]] <- 20
plot_mat(mat1, main="t[250]")
dev.off()

# generate gif
# library(animation)
# saveHTML({
#     for(i in seq(1, n_iter, round(n_iter/50, 0))) {
#         mat_t <- test_mat; mat_t[ranges_sim[[i]]] <- 10
#         plot_mat(mat_t)
#     }
# },title = "Simulated spread, rule=1", description = desc, verbose = FALSE)

## getting connectivity measures
# permitted_cells <- which(permitted_vec == 1)
# rule <- 1
# dat <- data.frame(x1 = rep(permitted_cells, (rule*2 + 1)^2 - 1),
#                   x2 = get_adjacent(permitted_cells, n_row, n_col, rule)) %>%
#     filter(x2 %in% permitted_cells)
# graph <- graph_from_data_frame(dat)
# members <- components(graph)$membership %>%
#     data.frame(membership = ., id_cell = as.integer(names(.)))
# members$id_cell[members$membership %in% members$membership[members$id_cell %in% start_cells]]
```
